Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-03-03 spie/_region_.tex
yet another draft

\message{ !name(paper.tex) !offset(395) }
\subsection{Image Occlusion}
\label{sec:imageOcclusion}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{occlusiondiagram.jpg}}
  \hspace{0.4cm} \subfloat[][]{\includegraphics[width=3in, height=1.8in]{occlusionimages.png}}
  \caption{(a) The green image contains texture that belongs to the red surface, which should not be projected onto the blue target surface. (b) Above, without geometry alignment, texture to the left of the red line would be removed, which would leave some erroneous texture projected onto our target surface. Below, after geometry alignment, the correct amount of texture is removed.}
  \label{fig:occlusion}
\end{figure}

In order to properly texture surfaces such as the one shown in Figure
\ref{fig:occlusion}, it is important to detect when parts of image
projections contain texture belonging to other surfaces and should
therefore not be used. This can be tested by recursively performing
ray-polygon intersection tests between camera locations and other
surfaces in a regularly spaced grid \cite{rayintersection}. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, the recursion stops. Where
there is a mixture of both, the rectangular region is subdivided into
four, and the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\message{ !name(paper.tex) !offset(791) }

>>>>

\message{ !name(paper.tex) !offset(435) }
\subsection{Image Occlusion}
\label{sec:imageOcclusion}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{occlusiondiagram.png}}
   \subfloat[][]{\includegraphics[width=3in,
    height=1.8in]{occlusionimagebad.png}}\\
  \hspace{3in} 
  \subfloat[][]{\includegraphics[width=3in,
    height=1.8in]{occlusionimagegood.png}}

  \caption{(a) The image from the camera in this diagram contains
    texture that belongs to the grey occluding surface, which should
    not be projected onto the orange target surface; (b) without
    geometry alignment, texture to the left of the red line would be
    removed, which would leave some erroneous texture projected onto
    our target surface; (c) after geometry alignment, the image is
    shifted, resulting in the correct amount of texture being
    removed.}
  \label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a grey
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and other surfaces \cite{rayintersection}. These intersection tests
are performed at the corners of a grid overlaid upon the target
surface. Where all four corners of a grid section are occluded,
texture is removed. Where one or more corners are occluded, the grid
is subdivided into four, and the process repeats. Occlusion checking
works entirely with geometry, so by ensuring that images match
geometry using \ref{sec:geometryAlignment}'s alignment procedure,
texture belonging to other surfaces is accurately removed, as seen in
Figure \ref{fig:occlusion}(b).

\message{ !name(paper.tex) !offset(852) }

### 2013-03-03 spie/paper.tex
yet another draft

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

\begin{figure}
  \centering
  \includegraphics[trim=0cm 3cm 0cm 0cm, clip=true,
  width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

@@

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar environments, and
rotational differences between camera poses
\cite{szeliski2006image}. Feature matching however, works best when
multiple unique visual references exist in the environment that can be
detected in multiple images. In contrast, indoor environments have a
high prevalence of bare surfaces, as well as repeating textures, such
as similar windows, doors, and wall-mounted decorations, that are
difficulte to disambiguate. This lack of reliable reference points
often results in errors when matching images together.

@@

While this procedure can be used to generate a texture, it is also
useful for obtaining a collection of images that can be used as a
starting point for our proposed texturing procedure, as explained in
Section \ref{sec:2dAlignment}. For our datasets, this process
generally selects around 10\% of all the images that could be used for
texturing the surface. This roughly 10:1 subsampling not only reduces
computational complexity of the remaining steps, but also selects the
most preomising images for the remaining steps.

@@

Our proposed 2D alignment procedure consists of three parts. First,
images are projected onto the surface and lines within these projected
images are detected. Images are then transformed such that these lines
match geometric lines composing the boundaries of the surface being
textured. Second, occlusion checks are performed to remove invalid
parts of each image for the target surface. Third, SIFT feature
matches are detected between pairs of images, and a weighted linear
least squares problem is solved in order to maximize all image and
geometry-based alignments. Each step will now be explained in detail.

@@

  \centering \subfloat[][]{\includegraphics[trim=0cm 3cm 0cm 0cm,
    clip=true, width=5in]{matchlines.jpg}}

@@

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. Thus, with 2
or more pairs, it is possible to obtain a fixed rotation and
translation, which are saved for usage in Section
\ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line and apply a rotation
and the minimal translation to match its lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. Finally, in the case
where there are no line pairs, we can still rotate images in order to
exploit patterns in indoor environments. For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to exterior walls corresponding to the edges of ceiling and
floor surfaces. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines. We use the RANSAC
framework, with an inlier difference of $0.5^\circ$ in order to
successfully ignore outliers \cite{fischler1981random}.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{occlusiondiagram.png}}
  \hspace{0.4cm} \subfloat[][]{\includegraphics[width=3in,
    height=1.8in]{occlusionimages.png}}
  \caption{(a) The image from the camera in this diagram contains
    texture that belongs to the grey occluding surface, which should
    not be projected onto the orange target surface; (b) without
    geometry alignment, texture to the left of the red line would be
    removed, which would leave some erroneous texture projected onto
    our target surface; (c) after geometry alignment, the image is
    shifted, resulting in the correct amount of texture being
    removed.}
  \label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a grey
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and other surfaces \cite{rayintersection}. These intersection tests
are performed at the corners of a grid overlaid upon the target
surface. Where all four corners of a grid section are occluded,
texture is removed. Where one or more corners are occluded, the grid
is subdivided into four, and the process repeats. Occlusion checking
works entirely with geometry, so by ensuring that images match
geometry using \ref{sec:geometryAlignment}'s alignment procedure,
texture belonging to other surfaces is accurately removed, as seen in
Figure \ref{fig:occlusion}(b).

@@

\section{Image Compositing}
\label{sec:imageCompositing}
In Section \ref{sec:mappingWithCaching}, we refine the tile-based
texturing approach from Section \ref{sec:simpleTextureMapping}, with
an added caching mechanism to reduce image boundaries. This method is
general and works well given arbitrary camera poses and surfaces such
as walls, floors, and ceilings. For special cases where images have
consistently perpendicular viewing angles to the surfaces under
consideration, such as walls, it is possible to develop an alternative
method in Section \ref{sec:shortestPath} to further reduce visual
artifacts.

>>>>

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{multirow}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

\begin{figure}
  \centering
  \includegraphics[width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

@@

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar environments, and
rotational differences between camera poses
\cite{szeliski2006image}. Feature matching however, works best when
multiple unique visual references exist in the environment that can be
detected in multiple images. In contrast, indoor environments have a
high prevalence of bare surfaces, as well as repeating textures, such
as similar windows, doors, and wall-mounted decorations, that are
difficult to disambiguate. This lack of reliable reference points
often results in errors when matching images together.

@@

While this procedure can be used to generate a texture, it is also
useful for obtaining a collection of images that can be used as a
starting point for our proposed texturing procedure, as explained in
Section \ref{sec:2dAlignment}. For our datasets, this process
generally selects around 10\% of all the images that could be used for
texturing the surface. This roughly 10:1 subsampling not only reduces
computational complexity of the remaining steps, but also selects the
most promising images for the remaining steps.

@@

Our proposed 2D alignment procedure consists of three parts, as shown
in the diagram in Figure \ref{fig:flowchart}. First, images are
projected onto the surface and lines within these projected images are
detected. Images are then transformed such that these lines match
geometric lines composing the boundaries of the surface being
textured. Second, occlusion checks are performed to remove invalid
parts of each image for the target surface. Third, SIFT feature
matches are detected between pairs of images, and a weighted linear
least squares problem is solved in order to maximize all image and
geometry-based alignments. Each step will now be explained in detail.

@@

  \centering \subfloat[][]{\includegraphics[trim=0cm 2cm 0cm 0cm,
    clip=true, width=5in]{matchLines.jpg}}

@@

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. This
corresponds to case (1) of Figure
\ref{fig:geometryAlignment}(c). Thus, with 2 or more pairs, it is
possible to obtain a fixed rotation and translation, which are saved
for usage in Section \ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line and apply a rotation
and the minimal translation to match its lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. This corresponds to
case (2) of Figure \ref{fig:geometryAlignment}(c). Finally, in the
case where there are no line pairs, we can still rotate images in
order to exploit patterns in indoor environments (case (3) of Figure
\ref{fig:geometryAlignment}(c)). For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to exterior walls corresponding to the edges of ceiling and
floor surfaces. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines. We use the RANSAC
framework, with an inlier difference of $0.5^\circ$ in order to
successfully ignore outliers \cite{fischler1981random}.

@@

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-1in}{\includegraphics[width=3in]{occlusiondiagram.png}}} &

  \begin{tabular}{c}
    \subfloat[][]{\includegraphics[trim=0cm 0cm 1.3cm 0cm,
    clip=true, width=3in,
      height=0.9in]{occlusionimagebad.png}} \\
    \subfloat[][]{\includegraphics[trim=1.3cm 0cm 0cm 0cm,
    clip=true, width=3in,
      height=0.9in]{occlusionimagegood.png}}
  \end{tabular}
\end{tabular}
\caption{(a) The image from the camera in this diagram contains
  texture that belongs to the gray occluding surface, which should not
  be projected onto the orange target surface; (b) without geometry
  alignment, texture to the left of the red line would be removed,
  which would leave some erroneous texture projected onto our target
  surface; (c) after geometry alignment, the image is shifted,
  resulting in the correct amount of texture being removed.}
\label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and other surfaces \cite{rayintersection}. These intersection tests
are performed at the corners of a grid overlaid upon the target
surface. Where all four corners of a grid section are occluded,
texture is removed. Where one or more corners are occluded, the grid
is subdivided into four, and the process repeats. Occlusion checking
works entirely with geometry, so by ensuring that images match
geometry using \ref{sec:geometryAlignment}'s alignment procedure,
texture belonging to other surfaces is accurately removed, as seen in
Figure \ref{fig:occlusion}(b).

@@

\section{Image Compositing}
\label{sec:imageCompositing}
In Section \ref{sec:mappingWithCaching}, we refine the tile-based
texturing approach from Section \ref{sec:simpleTextureMapping}, with
an added caching mechanism to reduce image boundaries. This method is
general and works well given arbitrary camera poses and surfaces such
as walls, floors, and ceilings. For special cases where images have
consistently perpendicular viewing angles to the surfaces under
consideration, such as walls, it is possible to develop an alternative
method in Section \ref{sec:shortestPath} to further reduce visual
artifacts. Both of these approaches are followed by a blending step in
order to produce final textures for each surface, as shown in Figure
\ref{fig:flowchart}.
